---
title: "Sequential Cosine Similarity Feature Grouping (SCSFG)"
author: "Ethan Ashby"
date: "12/16/2020"
output: pdf_document
---

###Package Load
```{r}
library(philentropy)
library(variantprobs)
library(tidyverse)
library(data.table)
library(gtools)
library(RColorBrewer)
library(gplots)
library(DescTools)
library(reticulate)
library(splus2R)
source("calc_minfo copy.R")
library(qgraph)
library(huge)
library(TCGAbiolinks)
library(utils)
`%!in%` = Negate(`%in%`)
```

###Load data

```{r}
data(tcga)
```


###Write function that calculates feature similarities and groups

My method begins by generating a set of dummy vectors that correspond to class attributes. For example, the dummy vector for the cancer type `ACC` is $(1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0)$. The user is provided the option of how many dummy vectors they would like to generate using the `dummy_exp` argument. `dummy_exp=1` yields $32$ class attribute vectors, each corresponding to a unique cancer type. `dummy_exp=2` yields $32+\frac{32*31}{2}$ class attribute vectors, corresponding to unique cancer types AND all unique pairs of cancer types.

There are 19020 genes mutated in the TCGA dataset, the vast majority of which are sparsely mutated. Finding a way to efficiently assess similarity between many features is a major challenge in information processing and retrieval. Fortunately, cosine similarity only considers the nonzero features, leading to its especially efficient performance on sparse data. The method offers an *a priori* mutation threshold filter, allowing the removal of extremely sparse signals (i.e. genes only mutated once across all samples). The cosine similarity between the reduced gene set and each of the class attribute vectors are calculated, generating lists of genes ranked in order of descending similarity to each particular class attribute vector.

Distinguishing sparse signals from noise is a very difficult task. To address this challenge, I introduce gene expression as a metafeature to guide the aggregation of mutation signals. Gene expression is a known determinant of mutational processes (via gene expression coupled repair), so genes with similar expression patterns will tend to display similar mutation processes. Using the `GEPIA2` API, I identfied lists of similarly expressed genes for metagene candidates in TCGA tumor samples. Genes that showed 

For each class attribute vector, a metagene is constructed by sequentially stepping through the ranked list and adding that feature's mutation data to the gene group. As each feature is added to the metagene, the normalized mutual information (NMI) between the mutation probabilities and cancer probabilities is calculated, to assess the tissue specificity mutation signal. This procedure is repeated until the NMI is locally maximized. The metagene that generates this locally maximized NMI is the final metagene.

```{r}

### Thank you Roland https://stackoverflow.com/questions/28368072/create-combinations-of-a-binary-vector
f.roland <- function(n, m) {
  ind <- combn(seq_len(n), m)
  ind <- t(ind) + (seq_len(ncol(ind)) - 1) * n
  res <- rep(0, nrow(ind) * n)
  res[ind] <- 1
  matrix(res, ncol = n, nrow = nrow(ind), byrow = TRUE)
}

#https://stats.stackexchange.com/questions/36309/how-do-i-find-peaks-in-a-dataset
argmax <- function(x, y, w=1, ...) {
  require(zoo)
  n <- length(y)
  y.smooth <- loess(y ~ x, ...)$fitted
  y.max <- rollapply(zoo(y.smooth), 2*w+1, max, align="center")
  delta <- y.max - y.smooth[-c(1:w, n+1-1:w)]
  i.max <- which(delta <= 0) + w
  list(x=x[i.max], i=i.max, y.hat=y.smooth)
}

test <- function(w, span) {
  peaks <- argmax(x, y, w=w, span=span)

  plot(x, y, cex=0.75, col="Gray", main=paste("w = ", w, ", span = ", span, sep=""))
  lines(x, peaks$y.hat,  lwd=2) #$
  y.min <- min(y)
  sapply(peaks$i, function(i) lines(c(x[i],x[i]), c(y.min, peaks$y.hat[i]), col="Red", lty=2))
  points(x[peaks$i], peaks$y.hat[peaks$i], col="Red", pch=19, cex=1.25)
}

SCSFG<-function(dummy_exp=1, heatmap=FALSE, bubbleplot=FALSE, n_view=100, heatmap_filename="heatmap_sim.jpeg", bubbleplot_filename="bubbleplot_metagenes.jpeg", mut_thresh=3){
  message("Loading data")
  tcga_nh <- data.table::setDT(tcga) %>% filter(MS=="Non-hypermutated")
  num_samps<-length(unique(tcga_nh$patient_id))

  tcga_v_f <- tcga_nh[
    # filter out all hypermutated tumors
    # & tumors with unknown cancer
    MS=="Non-hypermutated" & !is.na(Cancer_Code),
    ][, # variant frequencies by gene, cancer type
      # also save n_tumor per cancer
      n_tumor := length(unique(patient_id)),
      by = .(Cancer_Code)][,
                          .(v_f = length(unique(patient_id)), n_tumor = n_tumor[1]),
                          by = .(Variant, Hugo_Symbol, Cancer_Code)
                          ]
  
  #only return singletons
  tcga_v_f<- tcga_v_f %>% filter(v_f==1)

  cancer_nums<-tcga_v_f %>% dplyr::select(Cancer_Code, n_tumor) %>% distinct()
  cancer_nums<-cancer_nums[order(cancer_nums$Cancer_Code),]
  cancer_types<-cancer_nums$Cancer_Code %>% unique()

### matrix of mutation counts per tiss. type
  message("Calculating mutation frequencies")
  dat_c<-tcga_v_f           %>%
    dcast(
      Hugo_Symbol ~ Cancer_Code,
      value.var = "v_f",
      fill = 0
    ) %>%
    magrittr::set_rownames(.$Hugo_Symbol) %>%
    .[, Hugo_Symbol := NULL] %>%
    data.matrix()
  
  dat_c<-dat_c[rowSums(dat_c)>mut_thresh,]

### scale by number of tumors per cancer type
  dat<-matrix(ncol=32, nrow=dim(dat_c)[1])
  for(i in 1:dim(dat_c)[1]){
    dat[i,]<-dat_c[i,]/cancer_nums$n_tumor
  }
  rownames(dat)<-rownames(dat_c)
  colnames(dat)<-colnames(dat_c)

  message("Initializing dummy vectors")
  dummies<-matrix(ncol=32, nrow=0)
  if(dummy_exp==1){
    dummies<-f.roland(n=32, m=1)
  }
  if(dummy_exp>1){
      for(i in 1:dummy_exp){
      dummies<-rbind(dummies, f.roland(n=32, m=i))
    }
  }

  cols<-c()
  for(i in 1:dim(dummies)[1]){
    cols<-c(cols, paste(cancer_nums$Cancer_Code[which(dummies[i,]==1)], collapse="_"))
  }
  
  message("Calculating similarities... this may take awhile")
  mat_of_sims<-outer(FUN=Vectorize(function(X, Y){cosine_dist(dummies[Y,], dat[X,], testNA=FALSE)}), X=1:dim(dat)[1], Y=1:dim(dummies)[1])
  
  rownames(mat_of_sims)<-rownames(dat)
  colnames(mat_of_sims)<-cols

  if(heatmap==TRUE){
    message("Rendering Heatmap... this may take awhile")
    jpeg(file= heatmap_filename)
    heatmap.2(as.matrix(mat_of_sims[1:n_view,]), col=colorRampPalette(c("black", "white", "red")), Rowv=TRUE, Colv=FALSE, trace="none", dendrogram="none")
    message(paste("Heatmap outputted as:", heatmap_filename))
  }
  
  ###Figure out how to do grouping!
  #barplot(dat_c[which(mat_of_sims[,2]>0.95),]/cancer_nums$n_tumor)
  
  #one way using weights
  #vec<-c()
  #for (i in seq(0.5, 1, 0.005)){
  #vec<-c(vec, Gini(colSums(dat_c[which(mat_of_sims[,1]>i),]), n=cancer_nums$n_tumor))
  #}
  #plot(vec)
  
  #another way using counts/n_tumor
  #vec<-c()
  #for (i in seq(0.75, 1, 0.001)){
  #vec<-c(vec, Gini(colSums(dat_c[which(mat_of_sims[,1]>i),])/cancer_nums$n_tumor))
  #}
  #plot(vec, type="p+l")
  
  #d<-colSums(dat_c[which(mat_of_sims[,1]>vec[which(vec>0.96)[1]]),])
  
  #cancer_probs<-matrix(cancer_nums$n_tumor/sum(cancer_nums$n_tumor), ncol=32)
  #colnames(cancer_probs)<-cancer_nums$Cancer_Code
  
  #tmp<-d/cancer_nums$n_tumor
  #tmp<-tmp/sum(tmp)
  #tmp<-matrix(tmp, ncol=32)
  #colnames(tmp)<-cancer_nums$Cancer_Code
  
  ###Ridge budget on metafeatures (gene expression) of sparse genes to add to group
  ###
  message("Evaluating metagenes based on local maximal NMI")
  to_plot<-matrix(ncol=32, nrow=0)
  list_of_genes<-list()
  for(j in 1:32){
    print(j)
  ranked_genes=names(sort(mat_of_sims[,j], decreasing = TRUE))
  
  
  cancer_probs=cancer_nums$n_tumor/sum(cancer_nums$n_tumor)
  
  nmi<-c()
  for(i in 2:2000){
  
  #GT approx
  mat<- 1-exp(-colSums(dat_c[which(rownames(dat_c) %in% ranked_genes[1:i]),])/(cancer_nums$n_tumor+1))
  mat<-matrix(mat, ncol=32)
  colnames(mat)<-cancer_nums$Cancer_Code
  
  nmi <- suppressWarnings(c(nmi, calc_minfo(
  mat,
  cancer_prob=cancer_probs,
  binary_minfo = FALSE,
  normalize = TRUE
)))
  }
  x=1:500
  y=nmi[1:500]
  peaks<-argmax(x,y,w=2,span=0.1)
    
  to_plot<-rbind(to_plot, 1-exp(-colSums(dat_c[which(rownames(dat_c) %in% ranked_genes[1:peaks$x[1]]),])/(cancer_nums$n_tumor+1)))
  
  list_of_genes[[j]]<-dat_c[which(rownames(dat_c) %in% ranked_genes[1:peaks$x[1]]),] %>% rownames()
}
  bub_plot<-as.data.frame(to_plot)
  bub_plot<-melt(bub_plot)
  bub_plot$ID<-rep(paste("Metagene", cancer_nums$Cancer_Code), 32)
  
  #bubbleplot
    if(bubbleplot==TRUE){
    message("Rendering Bubbleplot")
    jpeg(file= bubbleplot_filename)
    bub_plot %>% ggplot()+geom_point(aes(x=variable, y=ID, size=value, color=ID))+theme_bw()+theme(axis.text.x=element_text(angle=90), legend.position="none")+xlab("")+ylab("")
    message(paste("Bubbleplot outputted as:", bubbleplot_filename))
    }
  return(list_of_genes)
}

metagene_list<-SCSFG(dummy_exp=1, heatmap=FALSE, bubbleplot=FALSE, n_view = 100, mut_thresh=3)

mgene1<-as.list(metagene_list[[1]])

dat_c[which(rownames(dat_c) %in% metagene_list[[6]]),]
```

```{r}
library(reticulate)
#gepia<-import("gepia")
#py_config()

Sys.setenv(RETICULATE_PYTHON = "/Users/ethanashby/Desktop/Thesis Work/python3/bin/python")
reticulate::py_config()
```

```{r}
repl_python()
import sys
import os
import gepia
#import pandas as pd

#import sys
#print(sys.version)
#import os
#sys.path.append('/Library/Frameworks/Python.framework/Versions/3.6/lib/python3.6/site-packages')
#import gepia
#print(1, end='')
#for p in sys.path:
    #print(p)

datasets=['ACC_Tumor', 'BLCA_Tumor', 'BRCA_Tumor', 'CESC_Tumor', 'CHOL_Tumor', 'COAD_Tumor', 'DLBC_Tumor', 'ESCA_Tumor', 'GBM_Tumor', 'HNSC_Tumor', 'KICH_Tumor', 'KIRC_Tumor', 'KIRP_Tumor', 'LAML_Tumor', 'LGG_Tumor', 'LIHC_Tumor', 'LUAD_Tumor', 'LUSC_Tumor', 'MESO_Tumor', 'OV_Tumor', 'PAAD_Tumor', 'PCPG_Tumor', 'PRAD_Tumor', 'READ_Tumor', 'SARC_Tumor', 'SKCM_Tumor', 'STAD_Tumor', 'TGCT_Tumor', 'THCA_Tumor', 'THYM_Tumor', 'UCEC_Tumor', 'UCS_Tumor']

def extractDigits(lst): 
    return [[el] for el in lst] 

metagene_list=r.metagene_list
for i in range(2, len(metagene_list)):
  #set metagene
  mgene=metagene_list[i]
  #create sim object
  sim=gepia.similar()
  #create output directory and set as output directory
  dir=['./gepiaResults/Mgene', str(i+1)]
  os.mkdir("".join(dir))
  dir1=['./gepiaResults/Mgene', str(i+1), "/"]
  sim.setOutDir("".join(dir1))

  #All tumor datasets takes way too long, just run on dataset of native metagene
  sim.setParam('dataset', [datasets[i]])

  genes=extractDigits(mgene)

  for gene in genes:
    sim.setParam('signature', gene)
    sim.query()
```

```{r}
folders<-paste("Mgene", 1:32, sep="")
filtered_mgenes<-list()
for (folder in folders){
  print(folder)
  files<-list.files(paste("gepiaResults/", folder, sep=""), pattern="*.tsv", full.names=TRUE)
  ldf <- lapply(files, read.table)
  gene_names<-gsub("_.*", "", gsub(paste("gepiaResults/", folder, "/", sep=""), "", files))
  #anno<-read.delim("custom.txt")

####Speed this up
  mat_of_intersect_lengths<-matrix(ncol=length(ldf), nrow=length(ldf))
  for(i in 1:length(ldf)){
    mgene1<-data.frame("Ensembl.gene.ID"=gsub("\\..*", "", ldf[[i]]$V1))
    #joined1<-left_join(mgene1, anno, by="Ensembl.gene.ID")
  
    for(j in 1:length(ldf)){
      mgene2<-data.frame("Ensembl.gene.ID"=gsub("\\..*", "", ldf[[j]]$V1))
      #joined2<-left_join(mgene2, anno, by="Ensembl.gene.ID")
    
      mat_of_intersect_lengths[i,j]<-length(intersect(na.omit(mgene1$Ensembl.gene.ID), na.omit(mgene2$Ensembl.gene.ID)))
    }
  }

  rownames(mat_of_intersect_lengths)<-gene_names
  colnames(mat_of_intersect_lengths)<-gene_names

  mat_of_sims<-mat_of_intersect_lengths/1000
  #for hclust
  #mat_of_dists<-1-mat_of_sims

  #must intersect w/ another gene's list at least 25% of time
  add_to_list<-c()
  for(i in 1:dim(mat_of_sims)[1]){
    if(any(mat_of_sims[i,-i]> 0.25)){add_to_list<-c(add_to_list, rownames(mat_of_sims)[i])}
  }
  filtered_mgenes[[folder]]<-add_to_list
}

#hc<-hclust(as.dist(mat_of_dists))
#plot(hc)
filtered_mgenes

to_plot<-matrix(ncol=32, nrow=0)
for (i in 1:32){
to_plot<-rbind(to_plot, 1-exp(-colSums(dat_c[which(rownames(dat_c) %in% filtered_mgenes[[i]]),])/(cancer_nums$n_tumor+1)))
}

bub_plot<-as.data.frame(to_plot)
bub_plot<-melt(bub_plot)
bub_plot$ID<-rep(paste("Metagene", cancer_nums$Cancer_Code), 32)

    bub_plot %>% ggplot()+geom_point(aes(x=variable, y=ID, size=value, color=ID))+theme_bw()+theme(axis.text.x=element_text(angle=90), legend.position="none")+xlab("")+ylab("")
```

### New procedure: rearranged, using TCGA biolinks

```{r}

foo<-function(){}

CosGLASSOFG<-function(dummy_exp=1, mut_thresh=0, sim_cutoff=0.7, nlambda=30, lambda.min.ratio=1E-6){
  #################################
  #Loading data
  #################################
  message("Loading data")
  tcga_nh <- data.table::setDT(tcga) %>% filter(MS=="Non-hypermutated")
  num_samps<-length(unique(tcga_nh$patient_id))

  tcga_v_f <- tcga_nh[
    # filter out all hypermutated tumors
    # & tumors with unknown cancer
    MS=="Non-hypermutated" & !is.na(Cancer_Code),
    ][, # variant frequencies by gene, cancer type
      # also save n_tumor per cancer
      n_tumor := length(unique(patient_id)),
      by = .(Cancer_Code)][,
                          .(v_f = length(unique(patient_id)), n_tumor = n_tumor[1]),
                          by = .(Variant, Hugo_Symbol, Cancer_Code)
                          ]
  
  #only return singletons
  tcga_v_f<- tcga_v_f %>% filter(v_f==1)

  cancer_nums<-tcga_v_f %>% dplyr::select(Cancer_Code, n_tumor) %>% distinct()
  cancer_nums<-cancer_nums[order(cancer_nums$Cancer_Code),]
  cancer_types<-cancer_nums$Cancer_Code %>% unique()

### matrix of mutation counts per tiss. type
  message("Calculating mutation frequencies")
  dat_c<-suppressMessages(tcga_v_f           %>%
    dcast(
      Hugo_Symbol ~ Cancer_Code,
      value.var = "v_f",
      fill = 0
    ) %>%
    magrittr::set_rownames(.$Hugo_Symbol) %>%
    .[, Hugo_Symbol := NULL] %>%
    data.matrix())
  
  dat_c<-dat_c[rowSums(dat_c)>mut_thresh,]

### scale by number of tumors per cancer type
  dat<-matrix(ncol=32, nrow=dim(dat_c)[1])
  for(i in 1:dim(dat_c)[1]){
    dat[i,]<-dat_c[i,]/cancer_nums$n_tumor
  }
  rownames(dat)<-rownames(dat_c)
  colnames(dat)<-colnames(dat_c)

  message("Initializing dummy vectors")
  dummies<-matrix(ncol=32, nrow=0)
  if(dummy_exp==1){
    dummies<-f.roland(n=32, m=1)
  }
  if(dummy_exp>1){
      for(i in 1:dummy_exp){
      dummies<-rbind(dummies, f.roland(n=32, m=i))
    }
  }

  cols<-c()
  for(i in 1:dim(dummies)[1]){
    cols<-c(cols, paste(cancer_nums$Cancer_Code[which(dummies[i,]==1)], collapse="_"))
  }
  
  message("Calculating similarities... this may take a second")
  mat_of_sims<-outer(FUN=Vectorize(function(X, Y){cosine_dist(dummies[Y,], dat[X,], testNA=FALSE)}), X=1:dim(dat)[1], Y=1:dim(dummies)[1])
  
  rownames(mat_of_sims)<-rownames(dat)
  colnames(mat_of_sims)<-cols

  #################
  #TCGA biolinks
  ################
  projects<-TCGAbiolinks::getGDCprojects()$project_id
  projects <- projects[grepl('^TCGA',projects,perl=T)]
  projects<- projects %>% sort()
  ##########
  #Removed TCGA-COAD, doing TCGA-READ
  projects<-projects[-24]
  projects[6]<-"TCGA-READ"

  #message("Evaluating metagenes based on local maximal NMI")
output_list<-list()
for(proj in projects){
    j=which(proj==projects)
    ranked_genes=names(sort(mat_of_sims[,j], decreasing = TRUE))
    sims<-sort(mat_of_sims[,j], decreasing = TRUE)
    ranked_genes=names(sims[sims>sim_cutoff])
    
    #Read in rda files if already downloaded
    if(paste(proj, ".rda", sep="") %in% list.files()){
      print(paste("Found data for: ", proj, "... reading it in now", sep=" "))
      file=paste(proj, ".rda", sep="")
      data<-suppressMessages(suppressWarnings(attach(file)))
      data<-data$data
    }
    
    #Generate rda files if not yet downloaded
    if(paste(proj, ".rda", sep="") %!in% list.files()){
      print(paste("Retrieving expression data for:", proj, sep=" "))
      
      if(proj %!in% c("TCGA-COAD", "TCGA-READ")){
      barcodes<-tcga_nh[paste("TCGA-", tcga_nh$Cancer_Code, sep="")==proj]$patient_id}
      if(proj %in% c("TCGA-COAD", "TCGA-READ")){
      barcodes<-tcga_nh[paste("TCGA-", tcga_nh$Cancer_Code, sep="")=="TCGA-COADREAD"]$patient_id}
      
      query <- GDCquery(project = proj,
                      data.category = "Gene expression",
                      data.type = "Gene expression quantification",
                      platform = "Illumina HiSeq", 
                      file.type  = "normalized_results",
                      experimental.strategy = "RNA-Seq",
                      barcode=barcodes,
                      legacy = TRUE)
      tryCatch(GDCdownload(query, method = "api", files.per.chunk = 20),
              error = function(e) GDCdownload(query, method = "client"))
      data<-GDCprepare(query, save=TRUE, save.filename=paste(proj, ".rda", sep=""))}

    message(paste("Calculating coexpression matrix"))
    norm_mat<-data@assays@data$normalized_count
    rownames(norm_mat)<-data@rowRanges$gene_id
    
    #subset for only genes in ranked list
    norm_mat<-norm_mat[which(rownames(norm_mat) %in% ranked_genes), ]
    
    #calculate covariance matrix and remove duplicates
    cov_mat<-cov(t(norm_mat))
    dup<-duplicated(rownames(cov_mat))
    if(any(dup)==TRUE){
    cov_mat<-cov_mat[c(-dup),c(-dup)]}
    if(any(dup)==FALSE){
    foo()
    }
    
    #calculate multiple glasso matrices
    message(paste("Graphical LASSO on covariance matrix of gene expressions"))
    glasso_mat<-huge::huge(cov_mat, method="glasso", nlambda=nlambda, lambda.min.ratio = lambda.min.ratio)
    
    message(paste("Retrieve connected gene modules and store as candidate gene groups"))
    indices<-which(glasso_mat$sparsity!=0)
    reg_genes<-list()
    #return regularized edge weights
    if(length(indices)==0){foo()}
    if(length(indices)>0){
    for(i in 1:length(indices)){
      index<-indices[i]
      adj<-glasso_mat$path[[index]]
      a<-rowSums(adj)
      reg_genes[[i]]<-rownames(cov_mat)[a>0]
    }}
    
    output_list[[length(output_list)+1]]<-reg_genes
}

return(output_list)
}
```


```{r}
#retrieve gene expression information using TCGAbiolinks
mgenes_schmorgasboard<-CosGLASSOFG(dummy_exp=1, mut_thresh=0, sim_cutoff=0.7, nlambda=30, lambda.min.ratio = 0.0001)
```

```{r}
nmi_eval_mgenes<-function(mgenes_schmorgasboard){
  #################################
  #Loading data
  #################################
  message("Loading data")
  tcga_nh <- data.table::setDT(tcga) %>% filter(MS=="Non-hypermutated")
  num_samps<-length(unique(tcga_nh$patient_id))

  tcga_v_f <- tcga_nh[
    # filter out all hypermutated tumors
    # & tumors with unknown cancer
    MS=="Non-hypermutated" & !is.na(Cancer_Code),
    ][, # variant frequencies by gene, cancer type
      # also save n_tumor per cancer
      n_tumor := length(unique(patient_id)),
      by = .(Cancer_Code)][,
                          .(v_f = length(unique(patient_id)), n_tumor = n_tumor[1]),
                          by = .(Variant, Hugo_Symbol, Cancer_Code)
                          ]
  
  #only return singletons
  tcga_v_f<- tcga_v_f %>% filter(v_f==1)

  cancer_nums<-tcga_v_f %>% dplyr::select(Cancer_Code, n_tumor) %>% distinct()
  cancer_nums<-cancer_nums[order(cancer_nums$Cancer_Code),]
  cancer_types<-cancer_nums$Cancer_Code %>% unique()

### matrix of mutation counts per tiss. type
  message("Calculating mutation frequencies")
  dat_c<-suppressMessages(tcga_v_f           %>%
    dcast(
      Hugo_Symbol ~ Cancer_Code,
      value.var = "v_f",
      fill = 0
    ) %>%
    magrittr::set_rownames(.$Hugo_Symbol) %>%
    .[, Hugo_Symbol := NULL] %>%
    data.matrix())
  
  dat_c<-dat_c[rowSums(dat_c)>mut_thresh,]
  
  ### scale by number of tumors per cancer type
  dat<-matrix(ncol=32, nrow=dim(dat_c)[1])
  for(i in 1:dim(dat_c)[1]){
    dat[i,]<-1-exp(-dat_c[i,]/cancer_nums$n_tumor)
  }
  rownames(dat)<-rownames(dat_c)
  colnames(dat)<-colnames(dat_c)
  
  cancer_probs<-cancer_nums$n_tumor/num_samps
  
#null 99th quantile gini
null_gini<-list()
for(i in 2:500){
    if(i %% 10 == 0){print(i)}
    tmp<-c()
    for(j in 1:1000){
    mat<-1-exp(-colSums(dat_c[sample(nrow(dat_c), i),])/(cancer_nums$n_tumor+1))
    mat<-matrix(mat, ncol=32)
    colnames(mat)<-cancer_nums$Cancer_Code
    tmp<-c(tmp, Gini(mat))
    }
    null_gini[[i]]<-tmp
}
  null_ref<-unlist(lapply(null_gini, quantile, probs=0.99))[-1]
  
  
#return mgenes with largest residual gini index
tmp<-list()
for(i in 1:length(mgenes_schmorgasboard)){
  gini<-c()
  for(j in 1:length(mgenes_schmorgasboard[[i]])){
      indices<-which(rownames(dat_c) %in% mgenes_schmorgasboard[[i]][[j]])
      if(length(indices)>1){
      mat<- 1-exp(-colSums(dat_c[indices,])/(cancer_nums$n_tumor+1))}
      if(length(indices)==1){
      mat<- 1-exp(-sum(dat_c[indices])/(cancer_nums$n_tumor+1))
      }
      mat<-matrix(mat, ncol=32)
      colnames(mat)<-cancer_nums$Cancer_Code
      
      gini <- c(gini, Gini(mat))
  }
  
  lens<-lapply(mgenes_schmorgasboard[[i]], length)
  #plot(2:250, null_ref, ylim=c(0,1), col="red", type="l")
  #points(lens, gini, col="blue")
  
  #par(pty="s")
  #plot(null_ref[unlist(lens)-1], gini, xlim=c(0,1), ylim=c(0,1), xlab="99th Null Gini Quantile", ylab="Metagene Gini")
  #text(null_ref[unlist(lens)-1], gini, labels=lens, cex=0.9, font=2, pos=3)
  #abline(a=0, b=1, col="red")
  
  resids<-unname(gini-null_ref[unlist(lens)-1])
  #plot(lens, resids)
  #text(lens, resids, labels=lens, cex=0.9, font=2, pos=3)
  #abline(h=max(resids, na.rm=TRUE))
  
  tmp<-c(tmp, c(list(mgenes_schmorgasboard[[i]][[which(resids==max(resids, na.rm=TRUE))[1]]]), max(resids, na.rm=TRUE)))
}
  tmp
}

tmp[c(TRUE, FALSE)]

barplot(unlist(tmp[c(FALSE, TRUE)]), names=cancer_nums$Cancer_Code, las=2, ylab="Increase in Gini Index compared to 99% quantile of size-matched random grouping")

```

```{r}
#Reconstruct to maximize Gini

gini_rerank<-function(mgenes_schmorgasboard){
  
  #################################
  #Loading data
  #################################
  message("Loading data")
  tcga_nh <- data.table::setDT(tcga) %>% filter(MS=="Non-hypermutated")
  num_samps<-length(unique(tcga_nh$patient_id))

  tcga_v_f <- tcga_nh[
    # filter out all hypermutated tumors
    # & tumors with unknown cancer
    MS=="Non-hypermutated" & !is.na(Cancer_Code),
    ][, # variant frequencies by gene, cancer type
      # also save n_tumor per cancer
      n_tumor := length(unique(patient_id)),
      by = .(Cancer_Code)][,
                          .(v_f = length(unique(patient_id)), n_tumor = n_tumor[1]),
                          by = .(Variant, Hugo_Symbol, Cancer_Code)
                          ]
  
  #only return singletons
  tcga_v_f<- tcga_v_f %>% filter(v_f==1)

  cancer_nums<-tcga_v_f %>% dplyr::select(Cancer_Code, n_tumor) %>% distinct()
  cancer_nums<-cancer_nums[order(cancer_nums$Cancer_Code),]
  cancer_types<-cancer_nums$Cancer_Code %>% unique()

### matrix of mutation counts per tiss. type
  message("Calculating mutation frequencies")
  dat_c<-suppressMessages(tcga_v_f           %>%
    dcast(
      Hugo_Symbol ~ Cancer_Code,
      value.var = "v_f",
      fill = 0
    ) %>%
    magrittr::set_rownames(.$Hugo_Symbol) %>%
    .[, Hugo_Symbol := NULL] %>%
    data.matrix())
  
  dat_c<-dat_c[rowSums(dat_c)>0,]
  
  ### scale by number of tumors per cancer type
  dat<-matrix(ncol=32, nrow=dim(dat_c)[1])
  for(i in 1:dim(dat_c)[1]){
    dat[i,]<-1-exp(-dat_c[i,]/cancer_nums$n_tumor)
  }
  rownames(dat)<-rownames(dat_c)
  colnames(dat)<-colnames(dat_c)
  
  cancer_probs<-cancer_nums$n_tumor/num_samps
  
  ############
  #dummies
  message("Initializing dummy vectors")
  dummies<-matrix(ncol=32, nrow=0)
  if(dummy_exp==1){
    dummies<-f.roland(n=32, m=1)
  }
  if(dummy_exp>1){
      for(i in 1:dummy_exp){
      dummies<-rbind(dummies, f.roland(n=32, m=i))
    }
  }

  
  #null 99th quantile gini
  message("Generating null gini index values... this may take a minute")
  null_gini<-list()
  for(i in 1:100){
      tmp<-c()
      for(j in 1:1000){
        if(i!=1){
          mat<-1-exp(-colSums(dat_c[sample(nrow(dat_c), i),])/(cancer_nums$n_tumor+1))
        }
        if(i==1){
          mat<-1-exp(-sum(dat_c[sample(nrow(dat_c), i),])/(cancer_nums$n_tumor+1))
        }
      mat<-matrix(mat, ncol=32)
      colnames(mat)<-cancer_nums$Cancer_Code
      tmp<-c(tmp, Gini(mat))
      }
      null_gini[[i]]<-tmp
  }
  
  #get null reference 99 quant gini values 
  null_ref<-unlist(lapply(null_gini, quantile, probs=0.99))[-1]
  
  #return mgenes with largest residual gini index
  message("Finding gene subset that contains maximal residual gini as proper subset")
  tmp<-list()
  for(i in 1:length(mgenes_schmorgasboard)){
  gini<-c()
  for(j in 1:length(mgenes_schmorgasboard[[i]])){
      indices<-which(rownames(dat_c) %in% mgenes_schmorgasboard[[i]][[j]])
      if(length(indices)>1){
      mat<- 1-exp(-colSums(dat_c[indices,])/(cancer_nums$n_tumor+1))}
      if(length(indices)==1){
      mat<- 1-exp(-sum(dat_c[indices])/(cancer_nums$n_tumor+1))
      }
      mat<-matrix(mat, ncol=32)
      colnames(mat)<-cancer_nums$Cancer_Code
      
      gini <- c(gini, Gini(mat))
  }
  
  lens<-lapply(mgenes_schmorgasboard[[i]], length)
  #plot(2:500, null_ref, ylim=c(0,1), col="red", type="l")
  #points(lens, gini, col="blue")
  
  #par(pty="s")
  #plot(null_ref[unlist(lens)-1], gini, xlim=c(0,1), ylim=c(0,1), xlab="99th Null Gini Quantile", ylab="Metagene Gini")
  #text(null_ref[unlist(lens)-1], gini, labels=lens, cex=0.9, font=2, pos=3)
  #abline(a=0, b=1, col="red")
  
  resids<-unname(gini-null_ref[unlist(lens)-1])
  #plot(lens, resids)
  #text(lens, resids, labels=lens, cex=0.9, font=2, pos=3)
  #abline(h=max(resids, na.rm=TRUE))
  
  ############
  #return one step up after the max
  if(which(resids==max(resids, na.rm=TRUE))==length(resids)){
  tmp<-c(tmp, c(list(mgenes_schmorgasboard[[i]][[which(resids==max(resids, na.rm=TRUE))[1]]], max(resids, na.rm=TRUE))))
  }
  if(which(resids==max(resids, na.rm=TRUE))<length(resids)){
  tmp<-c(tmp, c(list(mgenes_schmorgasboard[[i]][[which(resids==max(resids, na.rm=TRUE))[1]+1]], max(resids, na.rm=TRUE))))
  }
}

##########
#output reranked genes
message("Reranking genes and reconstructing metagene to maximize residual gini index")
output_list<-list()
  for(i in 1:32){
  rerank<-apply(dat[which(rownames(dat) %in% tmp[c(TRUE, FALSE)][[i]]),], MARGIN=1, cosine_dist, dummies[i,], testNA=FALSE) %>% sort(decreasing=TRUE)
  
  gini<-c()
  for(j in 1:length(names(rerank))){
  tmp_mgene<-names(rerank)[1:j]
  if(length(tmp_mgene)==1){
    mat<-1-exp(-sum(dat_c[which(rownames(dat_c) %in% tmp_mgene),])/(cancer_nums$n_tumor+1))
  }
  if(length(tmp_mgene)>1){
    mat<-1-exp(-colSums(dat_c[which(rownames(dat_c) %in% tmp_mgene),])/(cancer_nums$n_tumor+1))
  }
  
  mat<-matrix(mat, ncol=32)
  colnames(mat)<-cancer_nums$Cancer_Code
      
  gini <- c(gini, Gini(mat))
  }

  resids<-unname(gini-null_ref[1:length(rerank)])
  output_list[[i]]<-names(rerank)[1:which(resids==max(resids, na.rm=TRUE))]
  }

output_list
}

output_list<-gini_rerank(mgenes_schmorgasboard = mgenes_schmorgasboard)
```



```{r}
library(igraph)
library(diffusr)
HI<-read.table("HI-union.tsv")
custom <- read_delim("custom.csv", "\t", 
    escape_double = FALSE, trim_ws = TRUE)

binary_interacts<-left_join(left_join(HI, custom %>% select(`Approved symbol`,`Ensembl gene ID`), by=c("V1"="Ensembl gene ID")), custom %>% select(`Approved symbol`,`Ensembl gene ID`), by=c("V2"="Ensembl gene ID"))

g<-graph_from_data_frame(binary_interacts %>% dplyr::select(`Approved symbol.x`, `Approved symbol.y`) %>% slice(1:250), directed=FALSE)

adj<-as_adjacency_matrix(g)

adj<-adj[sort(rownames(adj)),sort(colnames(adj))]

adj<-hub.correction(as.matrix(adj))

h0<-apply(dat[which(rownames(dat) %in% rownames(adj)), ], MARGIN=1, FUN=Gini)
adj<-adj[which(rownames(adj) %in% names(h0)), which(colnames(adj) %in% names(h0))]

heat<-diffusr::heat.diffusion(h0=h0, adj, t=1000)

V(g)$heat<-heat
V(g)$color<-ifelse(V(g)$heat>quantile(heat, 0.99), "red", "white")
V(g)$size<-ifelse(V(g)$heat>quantile(heat, 0.99), 5, 1)
  
plot(g, vertex.label=NA)
```


Heat diffusion
LOCI (local correlation score for outlier detection)
